\section{Introduction}

Image processing, a fundamental task in computer vision and digital imaging, %
involves manipulating and enhancing digital images to extract useful information %
or to improve their visual quality. In many real-world applications, such as medical %
imaging, video processing, and remote sensing, the performance and efficiency of %
such algorithms are crucial for timely and accurate results. With the %
ever-increasing size and complexity of image data, there is a growing need for %
efficient parallel algorithms to handle the computational demands of this kinds %
of tasks.

The objective of this project was to explore and implement parallel image %
processing algorithms using shared and distributed memory paradigms and GPU accelerators. %
Specifically, we focused on parallelizing image filtering operations, which are %
commonly used for tasks such as noise reduction, edge detection, and image enhancement. %
We implemented and optimized a convolution-based image filtering %
algorithm to detect edges, which applies a filter kernel to each pixel in the %
input image to produce the corresponding pixel in the output image.

We began by developing a serial implementation of the algorithm %
and subsequently parallelized it using OpenMP for shared memory systems, MPI %
for distributed memory systems, and CUDA to harness the power of GPU cores. The parallel %
algorithms were designed to distribute the computational workload efficiently %
among available threads or processes, leveraging the capabilities of modern %
multi-core CPUs, distributed computing clusters and co-processors.

To ensure the correctness and effectiveness of the parallel implementations, %
appropriate boundary-handling strategies were employed to handle edge pixels %
during the convolution operation.\\
The project also analyzed performance and optimizations, including %
evaluating the speedup achieved by the parallel algorithms for different numbers %
of threads or processes.

Overall, this project aimed to provide a comprehensive understanding of parallel %
image filtering algorithms, their implementation, and their performance characteristics %
under different configurations. Through theoretical analysis and practical %
experimentation, we sought to gain insights into the scalability, %
efficiency, and adaptability of parallel image processing techniques.
